---
title: "simpleRegress"
author: "Kenroy Fuller"
date: "7/5/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

The bear neck is about 20 inches around the neck, and also 57 inches in length. The residual standard error is 5.376, and the R-squared is .7524 Rank # 4

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 



```{r}
MYCHEST <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(MYCHEST)
summary.lm(MYCHEST)
```

The residual standard error is 4.939 and its lower than the previous model which is 5.376 better. In addition, the R-squared is 0.791, which is higher and also better. So in ranking chest is above Neck so we can conclude chest is better than neck. Rank #5








```{r}
MYHEADWTH <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(MYHEADWTH)
summary.lm(MYHEADWTH)
```


As the third model, The residual standard error is 7.32, which is making it worse.In addition, the R-square is now 0.5409 which is also a decline from the previous model. comparing all three models, this is the one at the bottom of the list. Rank #6

















```{r}
MYWEIGHT<- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(MYWEIGHT)
summary.lm(MYWEIGHT)
```

This model is the comparison of the bear weight. The residual standard error is 5.432, and the adjusted R square is 0.7472.In this case it is higher than the last model which is worse, and the adjusted R-square is higher than the previous model which is better. So whereas, the residual standard error is worse, the adjusted R-square is better. at this point the bears chest has the highest adjusted R-square, which is the best. as far as the standard. Rank #3







```{r}
HEADLEN <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(HEADLEN)
summary.lm(HEADLEN)
```



 The Residual error is 4.235 fro the headlen while the R square is 0.8434 This one is ranked #1
 
 
 
 
 
 
 
 
 

```{r}
AGE <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(AGE)
summary.lm(AGE)
``` 

The Residual standard error is 7.511 while the adjusted R-square is 0.5073 Rank #2






Conclusion: In conclusion with all the rankings set in place you have to take into consideration the greatest residual which is headlen Ranked at #1 to dead last at the bear headwth which is ranked at #6 today we got to see as a class the statistical importance of the residual number, and also the R square!


